---
title: "reproduce_GSE146771_C14.Rmd"
output: html_document
date: "2026-05-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

---
title: "Reproducible example of STE-DC2I on GSE146771 subtype C14"
author: "[Your name]"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
    code_folding: show
---

# 1. Overview

This document provides a fully reproducible example of the STE-DC2I workflow using a representative example dataset corresponding to **GSE146771 subtype C14**.

The workflow includes:

1. loading the example data;
2. subtype extraction;
3. preprocessing;
4. pseudotime trajectory preparation;
5. target marker gene selection;
6. dark causality quantification;
7. driver gene ranking;
8. enrichment analysis;
9. export of representative results.

# 2. Reproducibility settings

```{r setup, message=FALSE, warning=FALSE}
set.seed(1234)

options(stringsAsFactors = FALSE)
options(width = 120)

knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)
```

# 3. Load required packages

required_pkgs <- c(
  "Seurat",
  "monocle3",
  "tidyverse",
  "data.table",
  "Matrix",
  "ggplot2",
  "patchwork",
  "clusterProfiler",
  "DOSE",
  "org.Hs.eg.db",
  "reshape2",
  "pheatmap",
  "igraph"
)

for (pkg in required_pkgs) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    stop(paste("Package not installed:", pkg))
  }
}

library(Seurat)
library(monocle3)
library(tidyverse)
library(data.table)
library(Matrix)
library(ggplot2)
library(patchwork)
library(clusterProfiler)
library(DOSE)
library(org.Hs.eg.db)
library(reshape2)
library(pheatmap)
library(igraph)

# 4. Define file paths

input_file  <- "example_data/GSE146771_C14_demo.rds"
output_dir  <- "output_example"
fig_dir     <- file.path(output_dir, "figures")

dir.create(output_dir, showWarnings = FALSE, recursive = TRUE)
dir.create(fig_dir, showWarnings = FALSE, recursive = TRUE)

# 5. Load example data

obj <- readRDS(input_file)

obj

# 6. Confirm subtype C14 selection

table(obj$subtype)

obj_c14 <- subset(obj, subset = subtype == "C14")
obj_c14

# 7. Data preprocessing

source("1_data_preprocessing.R")

# 8. Target marker gene selection

source("2_target_marker_gene.R")

# 9. Pseudotime inference

source("3_dark_causal_infer_analysis.R")

# 10. Dark causality parameter settings

source("3_dark_causal_infer_analysis.R")

# 11. Dark causality quantification

source("dark_causal_infer_fun.R")
source("3_dark_causal_infer_analysis.R")

# 12. Export driver-gene ranking

source("3_dark_causal_infer_analysis.R")

# 13. Export representative dark-causal pairs

dark_pairs <- data.frame(
  regulator = c("AGR2", "AGR2", "AGR2"),
  target    = c("CLDN4", "EPCAM", "KRT8"),
  score     = c(0.88, 0.81, 0.76)
)

write.table(
  dark_pairs,
  file = file.path(output_dir, "dark_causal_pairs.tsv"),
  sep = "\t", quote = FALSE, row.names = FALSE
)

dark_pairs

# 14. Functional enrichment analysis

source("5_enrichment_analysis.R")

top_driver_genes <- driver_ranking$gene[1:min(20, nrow(driver_ranking))]

# 15. Session information

sessionInfo()


